Robust Clustering for Ad Hoc Cognitive Radio Network

### 2017-04-09 journal-draft/ettdoc.tex
proofreading, modified minor mistakes

\newcommand{\bigO}{\ensuremath{\mathcal{O}}}% big-O notation/symbol
\usepackage{subfigure}
\usepackage{adjustbox}
\usepackage{multirow}
\usepackage{enumitem}
\usepackage{lipsum}
\setlist[itemize]{leftmargin=*}
\usepackage[referable]{threeparttablex}
\renewlist{tablenotes}{enumerate}{1}
\makeatletter
\setlist[tablenotes]{label=\tnote{\alph*},ref=\alph*,itemsep=\z@,topsep=\z@skip,partopsep=\z@skip,parsep=\z@,itemindent=\z@,labelindent=\tabcolsep,labelsep=.2em,leftmargin=*,align=left,before={\footnotesize}}
\makeatother

@@

In CRN, clusters are formed in the very beginning of the network operation, and re-formed periodically according to the dynamics of the CRN.
Each formed cluster has one or multiple unlicensed channels which are available for every CR node in the cluster.
The available unlicensed channels are referred to in the following of this paper as \textit{common licensed channels} (or common channels for short, which is abbreviated as CC).
Both payload and control overheads can be transmitted on the CCs.
%Out of the available CCs, there are always one or multiple channels which are used for the payload communication.
When one or several cluster members can not access one certain CC on which primary user activity is detected, the channel will be excluded from the set of CCs.
In particular, if that channel is being used for payload communication, the communication pair will stop and resume the transmission on another available CC.
The availability of CCs within a cluster defines the existence of that cluster, \ie no CCs are available means the corresponding cluster doesn't exist.
%As long as there is a CC available, the corresponding cluster can continue to exchange payload data.
In the context of CRN, the activity of primary users is usually unknown to the secondary users, then the primary users' activity is deemed as random.
In this case, the cluster which secures more CCs will anticipate a longer life expectancy, in other words, is more robust.
It is obvious that fewer secondary users in one cluster yield more CCs, but this contradicts to one of the motivations of clustering, \ie the cooperative decision making, where more users result in more accurate sensing result~\cite{Consensus_based_clustering12}.
Thus the number of small clusters, especially the \textit{singleton clusters}, \ie the cluster which has only one CR node, should be minimized.
As to the cluster which consists of many nodes, although the spectrum sensing is benefited, it usually leads to less common channels, which undermines the robustness of the cluster.
Thus the cluster robustness discussed in terms of number of CCs carries little meaning when the sizes of formed clusters are not given consideration.
Cluster size also plays a part in transmission power consumption, \ie cluster size affects the transmission power consumption when routing is conducted~\cite{clustering_globecom11, EnergyEfficientClusteringRouting_2015}.
In this paper we focus on the efficient formation of the clusters and the evaluation of the robustness of the formed clusters, but our scheme is also able to generate the clusters with the desired size, which gives the network designers more freedom when the power consumption is involved in the design of network applications.
%In this paper we will introduce a robust cluster structure for CRN, where the robustness means the ability of the clusters to sustain under the influence of the increasing activity of primary users, and the direct metric is the number of CCs in the clusters.

@@

Different clustering strategies have been proposed for wireless networks of different type.
As there is no restriction on the nodes to access the spectrum in wireless ad-hoc, mesh networks and sensor networks, the objective of clustering is to decrease the transmission power consumption~\cite{Kawadia03}, to improve the routing performance~\cite{clustering_mesh_globecom2010}, or to imrove the network lifetime and coverage~\cite{Abbasi_survey_07}.
%Various clustering schemes are proposed to target different aspects in cognitive radio networks.
%Work~\cite{Consensus_based_clustering12} improves spectrum sensing ability by grouping the CR users with potentially best detection performance into the same cluster.
%Clustering scheme~\cite{clustering_globecom11} obtains the best cluster size which minimizes power consumption caused by communication within and among clusters.
%\cite{clustering_globecom11} proposes clustering strategy in cognitive radio network, which looks into the relationship between cluster size and power consumption and accordingly controlling the cluster size to decrease power consumption.
With regard to forming clusters in CRN, deciding on the common channel within each cluster is the foremost question to answer.
\cite{Zhao07, Chen07,Affinity_clustering_09icccn} propose the clustering schemes and make sure every cluster possesses at least one CC.
Clustering scheme~\cite{Consensus_based_clustering12} looks for the network partition which improves the accuracy of spectrum sensing with the cluster structure.
%Clustering scheme~\cite{clustering_globecom11} obtains the best cluster size which minimizes power consumption caused by communication within and among clusters.
\cite{TWC2012_cooperative_communication} forms the clusters by deciding on the cluster heads, where the transmit power for the long-haul transmission between the cluster heads is minimized.
\cite{clustering_globecom11} proposes a cluster structure which promises energy efficiency.
\cite{cluster_EW10} proposes strategy on how to decide on the CCs and access the multiple CCs within clusters.
An event-driven clustering scheme is proposed for cognitive radio sensor network in \cite{Ozger_cluster_crsn_13}.
No one among the above mentioned schemes provides the robustness to the formed clusters against primary users.

@@

Second, this paper proposes size control mechanism, which solves the problem of devergence of clusters sizes in ~\cite{Li11_ROSS} and~\cite{LIU_TMC11_2}.

@@

We give the definition of cluster in CRN as follows. 
A cluster $C$ is a set of secondary nodes which possess the same set of common channels.
In particular, a cluster consists of a cluster head $h(C)$ and a number of cluster members, and the cluster head is able to communicate with any cluster member directly.
A cluster can be formed only by the cluster head.
%$\text{Nb}(i)$ denotes node $i$'s neighborhood which consists of all its one hop neighbors.
%For any cluster member $i \in C$, $i \in \text{Nb} (h_C) $ holds.
The size of $C$ is denoted by $|C|$.
When the cluster head of a cluster is $i$, we denote that cluster by $C(i)$.
$K(C)$ denotes the set of CCs in cluster $C$, $ K(C) = \Cap_{i\in C} K_i$.
%Clustering is performed periodically, because the secondary users are mobile and the primary users change their operation channels, thus the channel availability on secondary users changes accordingly.
The notations used in the system model are listed in Table~\ref{tab1}.
\begin{table}[h!]
\caption{Notations}
\label{tab1}
\centering
\begin{tabular}{llr}
\toprule
%\multicolumn{2}{c}{Item} \\
%\cmidrule(r){1-2}
Symbol & Description \\
\midrule
$\mathcal{N}$  & set of CR users in a CRN\\
$N$ & number of CR users in a CRN, $N=|\mathcal{N}|$\\
$\mathcal{K}$	& set of licensed channels\\
$k(i)$ & the working channel of user $i$\\
$\text{Nb}(i)$ & the neighborhood of CR node $i$    \\
$C(i)$ & a cluster whose cluster head is $i$  \\
%$k_i$   & set of available channels at CR node $i$  \\
$K_i$   & the set of available channels at CR node $i$  \\
$K(C(i))$   & the set of available CCs of cluster $C(i)$ \\
%$h(C)$ & cluster head of a cluster $C$\\
$h(C)$ & the cluster head of a cluster C\\
%$\text{CH}$ & cluster head\\
%$\text{CH}$ & cluster head\\
$\delta$ & the cluster size which is preferred\\
$S_i$ & a set of claiming clusters, each of which includes \\
& debatable node $i$ after phase I\\
$d_i$  & individual connectivity degree of CR node $i$\\
$g_i$  & neighborhood connectivity degree of CR node $i$\\
$f(C)$ & the number of CCs of a cluster $C$, which is used \\
& in the problem description\\
% $\mathcal{G}$ & a collection of some possible clusters in $\mathcal{N}$\\
 $\mathcal{S}$ & the collection of all the possible clusters in $\mathcal{N}$\\
 $C_i$  & the $i$-th cluster in $\mathcal{S}$ \\
 $|C_i|$ & size of the cluster $C_i$\\
 $|K(C_i)|$ & the number of CCs of cluster $C_i$\\
 $n$ & the number of debatable nodes\\
 $m$ & the number of claiming cluster heads\\
\bottomrule
\end{tabular}
\end{table}

@@

As to a cognitive radio network, where the set of CR nodes is $\mathcal{N}$.
Each cluster complies with the definition in~\ref{sec:model}, whose number of CCs in a cluster $C$ is $f(C)$.
The robust clustering problem is to decide the set of clusters $\mathcal{T}$, where 
\begin{enumerate}
\setlength{\itemindent}{-.17in}
\item the sum of the $f(C) is the maximal$, where $C\in \mathcal{T}$ 
\item the intersection of any two clusters in $\mathcal{T}$ is an empty set
\item the union of clusters in $\mathcal{T}$ is $\mathcal{N}$
\item the cluster sizes should fall in the scope $\langle\delta_1, \delta_2\rangle$, where $\delta, \delta_1, \delta_2\in \mathbb{Z}^+$, $\delta_1 \leq \delta\leq \delta_2$, and $\delta$ is decided according to the application to be implemented.
\end{enumerate}

@@

Assume we have obtained the set of $\mathcal{S}$ which contains all the possible clusters in $\mathcal{N}$, \ie $\mathcal{S}=\{C_1, C_2,\ldots,C_i, \ldots, C_{|\mathcal{S}|}\}$ \footnote{The subscript $i$ means the $i$-th cluster in $\mathcal{S}$.} and there is $\bigcup_{1\leq i \leq |\mathcal{S}|} C_i = \mathcal{N}$.
The decision version of this problem is to determine whether there is a non-empty set $\mathcal{S}'\subseteq \mathcal{S}$, so that $\sum_{C_i\in \mathcal{S}'} f(C_i) \geqslant \lambda$ where $\lambda$ is a real number.
We have the following theorem on the complexity of this problem.

@@

To formulate the debatable nodes' membership clarification into the desired congestion game, we reexamine this process from a different (or opposite) perspective. 
From the new perspective, the debatable nodes are not included in any cluster, and they are looking for the proper cluster to join.
% in other words, their claiming clusters become clusters which are beside them. 
%Now for the debatable nodes, the previous problem of deciding which clusters to leave becomes a new problem that which cluster to join.
A debatable node $i$ chooses one cluster $C$ out of $S_i$ to join if the decrease of CCs in cluster $C$ is the smaller than any other cluster in $S_i$.
The decrease of CCs in cluster $C$ is $\sum_{C\in S_i}\Delta\vert K(C) \vert=\sum_{C\in S_i}({\vert K(C) \vert-\vert K(C\cup i) \vert})$.
The interaction between the debatable nodes and the claiming clusters is shown in Figure~\ref{debatable_nodes_claiming_cluster}.
%The concern on convergence appears again as we have discussed in the previous subsubsection.
%We give proof on convergence under game theoretic framework.
\begin{figure}[ht!]
  \centering
  \includegraphics[width=0.25\linewidth]{singletongame_png.png}
  \caption{Debatable nodes and claiming clusters}
  \label{debatable_nodes_claiming_cluster}
\end{figure}

@@

In fact, the upper bound of the number of steps which are actually involved in this process is much smaller than $N^3$.%, as both $n$ and $m$ are considerably smaller than $N$.
The percentage of debatable nodes in the network is illustrated in Figure~\ref{percentage_overlapping_node}, which is between 10\% to 60\% of the total number of CR nodes in the network.
In the other hand, the number of clusters heads is dependent on the network density and the CR node's transmission range as mentioned in Section~\ref{phaseI}.
The simulation in \cite{robust_clustering_arxiv} shows that the cluster heads are only 3.4\% to 20\% of the total CR nodes.

@@

\item \textbf{Robustness of the clusters against newly added PUs.}
We increase the number of PUs to challenge the non-singleton clusters, and count the number of the unclustered CR nodes.
This metric directly indicates the robustness of clusters from a more practical point of view, \ie as to the clusters formed for a given CRN and spectrum availability, how many CR nodes can still make use of the clusters when the spectrum availability decreases.
%robustness of clusters, as this metric directly shows how many nodes can make use of the cluster structure.
%When we investigate the performance with moderate and vigorous intensity of primary users' activities, 
%All the robust clustering schemes are proposed for a given CRN, when the available spectrum becomes scarce when the primary users' operation becomes more intense, the metric can illustrate the robustness of clusters by showing the unclustered CR nodes.
%This metric is closely related with cluster robustness, \ie the less CR nodes which are not  nodes means more CR nodes are included into non-singleton clusters which survive the primary users' influence.
%When we vary the intensity of primary users' activity, \eg from low to medium level by increasing the number of primary users, 
%The drawback of this metric is, it doesn't indicate the robustness of the non-singleton clusters when the primary users conquer more spectrum.

@@

\subsubsection{Average number of CCs in Non-singleton Clusters}
\label{ccc_20}
%We first have a look at the average number of common channels per non-singleton cluster.
% shows the average number of the CCs of non-singleton clusters,
%as the singleton clusters (in other words unclustered nodes) don't execute any functionalities of clusters, which are has be discussed in Section~\ref{intro}.
From Figure~\ref{ccc_per_nonsingleton}, we can see the centralized schemes outperform the distributed schemes.
Among the distributed schemes, SOC achieves the most CCs.
The reason is, SOC is liable to group the neighboring CRs which share the most abundant spectrum together, no matter how many of them are there, thus the number of CC of the formed clusters is higher.
%We have discussed the flaw of this metric as it doesn't convey the number of unclustered CR nodes, in fact, 
On the other hand, SOC generates the most unclustered CRs.
As to the variants of ROSS, we notice that the greedy mechanism increases CCs in non-singleton clusters significantly.
%, this is due to the procedure that debatable nodes greedily look for better affiliation to improve the number of CCs.
%We also notice that the size control feature doesn't affect the number of CCs for both ROSS-DGA and ROSS-DFA.
%Size control mechanism converts the large clusters into small ones, but meanwhile clusters with the desired sizes have to be made when forming smaller clusters is possible.
%, which can also be observed in the large scale CRN in Section~\ref{largeScaleCRN}.

@@

\begin{figure*}[th]
\begin{multicols}{3}
    \includegraphics[width=\linewidth]{ccc_20.pdf}\par\caption{Number of common channels of non-singleton clusters}\label{ccc_per_nonsingleton}
    \includegraphics[width=\linewidth]{cdf_clusterSize_20.pdf}\par\caption{Cumulative distribution of CRs residing in clusters with different sizes}\label{size_control}    
    \includegraphics[width=\linewidth]{survival_rate_20.pdf}\par\caption{Number of unclustered CRs with decreasing spectrum availability}\label{singleton_clusters}
\end{multicols}
\caption{Comparison between the distributed and centralized clustering schemes ($N$ = 20)}
\label{compare_dis_centralized}
\end{figure*}

@@

\subsubsection{Robustness of the clusters against newly added PUs}
%When the number of PUs in CRN increases, or their operation becomes more intensive, some clusters don't seize any CCs any more, so that the cluster members and the cluster heads become unclustered, or singleton clusters.
%If there is no common channels available any more because of the new added PUs, the cluster is regarded as destroyed and the former cluster member CRs become unclustered CRs or in other words singleton clusters.
%We investigate the number of singleton clusters with primary users whose intensity of activities are varying.
%We will obtain two observations by examining this metric.
%First, we will know how many CR nodes are unclusterred by applying the robust clustering schemes in a given CRN.
In this part of simulation, we put PUs sequentially into CRN to decrease the available spectrum.
10 PUs are in the network in the beginning, then extra 19 batches of PUs are added sequentially, where each batch includes 5 PUs. 
%
Figure~\ref{singleton_clusters} shows certain clusters can not maintain and the number of unclustered CR nodes grows when the number of PUs increases.
The centralized scheme with desired size of 2 generates the most robust clusters, meanwhile, SOC results in the most vulnerable clusters.
The centralized scheme with desired size of 3 doesn't outperform the variants of ROSS, because pursuing cluster size prevents forming the the clusters with more CCs.
In contrary, the variants of ROSS generate some smaller clusters which are more likely to maintain when there are more PUs.

@@

\begin{figure}[ht!]
  \centering
  \includegraphics[width=.7\linewidth]{ccc_large_scale_color.pdf}
  \caption{Number of common channels of non-singleton clusters.}
  \label{ccc_large_scale}
\end{figure}

@@

\subsubsection{Robustness of the clusters against increased PUs}
We increase the primary users' activity by importing 20 batches of PUs sequentially in the CRN, where each batch includes 10 PUs.

@@

\begin{figure*}[t]
\begin{multicols}{3}
    \includegraphics[width=\linewidth]{survival_rate_100_edge50.pdf}\par\caption{100 CRs}\label{singleton_clusters_100}
    \includegraphics[width=\linewidth]{survival_rate_200_edge50.pdf}\par\caption{200 CRs}\label{singleton_clusters_200}
    \includegraphics[width=\linewidth]{survival_rate_300_edge50.pdf}\par\caption{300 CRs}\label{singleton_clusters_300}
\end{multicols}\caption{Percentage of CR nodes which are not included in any non-singleton clusters}
\label{unclustered_100_200_300}
\end{figure*}

@@

\begin{figure*}[t]
\begin{multicols}{3}
    \includegraphics[width=\linewidth]{cdf_clusterSize_100.pdf}\par\caption{100 CRs, 30 PUs in network}\label{cdf_clusterSize_100}
    \includegraphics[width=\linewidth]{cdf_clusterSize_200.pdf}\par\caption{200 CRs, 30 PUs in network}\label{cdf_clusterSize_200}
    \includegraphics[width=\linewidth]{cdf_clusterSize_300.pdf}\par\caption{300 CRs, 30 PUs in network}\label{cdf_clusterSize_300}
\end{multicols}
\caption{Cumulative distribution of CRs residing in clusters with different sizes}
\label{cdf_100_200_300}
\end{figure*}

>>>>

\newcommand{\bigO}{\ensuremath{\mathcal{O}}}% big-O notation/symbol
\usepackage{subfigure}
\usepackage{adjustbox}
\usepackage{multirow}
\usepackage{enumitem}
\usepackage{lipsum}
\setlist[itemize]{leftmargin=*}
\setlist[enumerate]{wide=\parindent}
\usepackage[referable]{threeparttablex}
\renewlist{tablenotes}{enumerate}{1}
\makeatletter
\setlist[tablenotes]{label=\tnote{\alph*},ref=\alph*,itemsep=\z@,topsep=\z@skip,partopsep=\z@skip,parsep=\z@,itemindent=\z@,labelindent=\tabcolsep,labelsep=.2em,leftmargin=*,align=left,before={\footnotesize}}
\makeatother

@@

In CRN, clusters are formed in the very beginning of the network operation, and re-formed periodically according to the dynamics of the CRN.
Each formed cluster has one or multiple unlicensed channels which are available for every CR node in the cluster.
The available unlicensed channels are referred to in the following of this paper as \textit{common licensed channels} (or common channels for short, which is abbreviated as CC).
Both payload and control overheads can be transmitted on the CCs.
%Out of the available CCs, there are always one or multiple channels which are used for the payload communication.
When one or several cluster members can not access one certain CC on which primary user activity is detected, the channel will be excluded from the set of CCs.
In particular, if that channel is being used for payload communication, the communication pair will stop and resume the transmission on another available CC.
The availability of CCs within a cluster defines the existence of that cluster, \ie no CCs are available means the corresponding cluster doesn't exist.
%As long as there is a CC available, the corresponding cluster can continue to exchange payload data.
In the context of CRN, the activity of primary users is usually unknown to the secondary users, then the primary users' activity is deemed as random.
In this case, the cluster which secures more CCs will anticipate a longer life expectancy, in other words, is more robust.
It is obvious that fewer secondary users in one cluster yield more CCs, but this contradicts to one of the motivations of clustering, \ie the cooperative decision making, where more users result in more accurate sensing result~\cite{Consensus_based_clustering12}.
Thus the number of small clusters, especially the \textit{singleton clusters}, \ie the cluster which has only one CR node, should be minimized.
As to the cluster which consists of many nodes, although the spectrum sensing is benefited, it usually leads to less common channels, which undermines the robustness of the cluster.
Thus the cluster robustness discussed in terms of number of CCs carries little meaning when the sizes of formed clusters are not given consideration.
Cluster size also plays a part in transmission power consumption, \ie cluster size affects the transmission power consumption when routing is conducted~\cite{clustering_globecom11, EnergyEfficientClusteringRouting_2015}.
In this paper we will analyze the robust clustering and propose efficient solutions.
Our schemes generate the clusters which demonstrate both the robustness against the increasing primary activity and the compliance with the desired cluster sizes.
Hence our schemes provide the network designers with more freedom when the power consumption is involved in the design of network applications.
%In this paper we will introduce a robust cluster structure for CRN, where the robustness means the ability of the clusters to sustain under the influence of the increasing activity of primary users, and the direct metric is the number of CCs in the clusters.

@@

Different clustering strategies have been proposed for wireless networks of different types.
As there is no restriction on the nodes to access the spectrum in wireless ad-hoc, mesh networks and sensor networks, the objective of clustering is to decrease the transmission power consumption~\cite{Kawadia03}, to improve the routing performance~\cite{clustering_mesh_globecom2010}, or to imrove the network lifetime and coverage~\cite{Abbasi_survey_07}.
%Various clustering schemes are proposed to target different aspects in cognitive radio networks.
%Work~\cite{Consensus_based_clustering12} improves spectrum sensing ability by grouping the CR users with potentially best detection performance into the same cluster.
%Clustering scheme~\cite{clustering_globecom11} obtains the best cluster size which minimizes power consumption caused by communication within and among clusters.
%\cite{clustering_globecom11} proposes clustering strategy in cognitive radio network, which looks into the relationship between cluster size and power consumption and accordingly controlling the cluster size to decrease power consumption.
With regard to forming clusters in CRN, deciding on the common channel within each cluster is the foremost question to answer.
\cite{Zhao07, Chen07,Affinity_clustering_09icccn} propose the clustering schemes and make sure every cluster possesses at least one CC.
Clustering scheme~\cite{Consensus_based_clustering12} looks for the network partition which improves the accuracy of spectrum sensing with the cluster structure.
%Clustering scheme~\cite{clustering_globecom11} obtains the best cluster size which minimizes power consumption caused by communication within and among clusters.
\cite{TWC2012_cooperative_communication} forms the clusters by deciding on the cluster heads, where the transmit power for the long-haul transmission between the cluster heads is minimized.
\cite{clustering_globecom11} proposes a cluster structure which promises energy efficiency.
\cite{cluster_EW10} proposes strategy on how to decide on the CCs and access the multiple CCs within clusters.
An event-driven clustering scheme is proposed for cognitive radio sensor network in \cite{Ozger_cluster_crsn_13}.
No one among the above mentioned schemes provides the robustness to the formed clusters against primary users.

@@

Second, this paper proposes size control mechanism, which solves the problem of devergence of clusters sizes in~\cite{Li11_ROSS} and~\cite{LIU_TMC11_2}.

@@

We give the definition of cluster in CRN as follows. 
A cluster $C$ is a set of secondary nodes, and there is one set of common channels which are available for each node.
In particular, a cluster consists of a cluster head $h(C)$ and a number of cluster members, and the cluster head is able to communicate with any cluster member directly.
A cluster can be formed only by the cluster head.
%$\text{Nb}(i)$ denotes node $i$'s neighborhood which consists of all its one hop neighbors.
%For any cluster member $i \in C$, $i \in \text{Nb} (h_C) $ holds.
The size of $C$ is denoted by $|C|$.
When the cluster head of a cluster is $i$, we denote that cluster by $C(i)$.
$K(C)$ denotes the set of CCs in cluster $C$, $ K(C) = \bigcap_{i\in C} K_i$.
%Clustering is performed periodically, because the secondary users are mobile and the primary users change their operation channels, thus the channel availability on secondary users changes accordingly.
The notations used in the system model are listed in Table~\ref{tab1}.
\begin{table}[h!]
\caption{Notations}
\label{tab1}
\centering
\begin{tabular}{llr}
\toprule
%\multicolumn{2}{c}{Item} \\
%\cmidrule(r){1-2}
Symbol & Description \\
\midrule
$\mathcal{N}$  & set of CR users in a CRN\\
$N$ & number of CR users in a CRN, $N=|\mathcal{N}|$\\
$\mathcal{K}$	& set of licensed channels\\
$k(i)$ & the working channel of user $i$\\
$\text{Nb}(i)$ & the neighborhood of CR node $i$    \\
$C(i)$ & a cluster whose cluster head is $i$  \\
%$k_i$   & set of available channels at CR node $i$  \\
$K_i$   & the set of available channels at CR node $i$  \\
$K(C(i))$   & the set of available CCs of cluster $C(i)$ \\
%$h(C)$ & cluster head of a cluster $C$\\
$h(C)$ & the cluster head of a cluster C\\
%$\text{CH}$ & cluster head\\
%$\text{CH}$ & cluster head\\
$\delta$ & the cluster size which is preferred\\
$S_i$ & a set of claiming clusters, each of which includes \\
& debatable node $i$ after phase I\\
$d_i$  & individual connectivity degree of CR node $i$\\
$g_i$  & neighborhood connectivity degree of CR node $i$\\
$f(C)$ & the number of CCs of a cluster $C$, which is used \\
& in the problem description\\
% $\mathcal{G}$ & a collection of some possible clusters in $\mathcal{N}$\\
 $\mathcal{S}$ & the collection of all the possible clusters in $\mathcal{N}$\\
 $C_i$  & the $i$-th cluster in $\mathcal{S}$ \\
 $|C_i|$ & size of the cluster $C_i$\\
 $|K(C_i)|$ & the number of CCs of cluster $C_i$\\
 $n$ & the number of debatable nodes\\
 $m$ & the number of claiming cluster heads\\
\bottomrule
\end{tabular}
\end{table}

@@

As to a cognitive radio network, where the set of CR nodes is $\mathcal{N}$.
%Each cluster complies with the definition in , whose number of CCs in a cluster $C$ is $f(C)$.
The robust clustering problem is to decide the set of clusters $\mathcal{T}$, where 
\begin{enumerate}
\setlength{\itemindent}{.05in}
\item the sum of the $f(C)$ is the maximal, where $C\in \mathcal{T}$ 
\item the intersection of any two clusters in $\mathcal{T}$ is an empty set
\item the union of clusters in $\mathcal{T}$ is $\mathcal{N}$
\item the cluster sizes should fall in the scope $\langle\delta_1, \delta_2\rangle$, where $\delta, \delta_1, \delta_2\in \mathbb{Z}^+$, $\delta_1 \leq \delta\leq \delta_2$, and $\delta$ is decided according to the application to be implemented in the CRN.
\end{enumerate}

@@

Assume we have obtained the set of $\mathcal{S}$ which contains all the clusters in $\mathcal{N}$, \ie $\mathcal{S}=\{C_1, C_2,\ldots,C_i, \ldots, C_{|\mathcal{S}|}\}$ \footnote{The subscript $i$ means the $i$-th cluster in $\mathcal{S}$.} and there is $\bigcup_{1\leq i \leq |\mathcal{S}|} C_i = \mathcal{N}$.
The decision version of this problem is to determine whether there is a non-empty set $\mathcal{S}'\subseteq \mathcal{S}$, so that $\sum_{C_i\in \mathcal{S}'} f(C_i) \geqslant \lambda$ where $\lambda$ is a real number.
We have the following theorem on the complexity of this problem.

@@

To formulate the debatable nodes' membership clarification into the desired congestion game, we reexamine this process from a different/opposite perspective. 
From the new perspective, the debatable nodes are not included in any cluster and they need to decide on one cluster to join.
When a debatable node $i$ join one cluster $C$, the decrease of CCs in cluster $C$ is $\sum_{C\in S_i}\Delta\vert K(C) \vert=\sum_{C\in S_i}({\vert K(C) \vert-\vert K(C\cup i) \vert})$.
Then, node $i$ chooses the cluster $C$, where the decrease of CCs in cluster $C$ is smaller than the decrease if $i$ would have joined any other claiming cluster in $S_i$.
The relation between the debatable nodes and the claiming clusters is shown in Figure~\ref{debatable_nodes_claiming_cluster}.
%The concern on convergence appears again as we have discussed in the previous subsubsection.
%We give proof on convergence under game theoretic framework.
\begin{figure}[ht!]
  \centering
  \includegraphics[width=0.25\linewidth]{singletongame_png.png}
  \caption{Debatable nodes and claiming clusters}
  \label{debatable_nodes_claiming_cluster}
\end{figure}

@@

In fact, the upper bound for the number of steps which are actually involved in this process is much smaller than $N^3$.
The percentage of debatable nodes in the network is shown in Figure~\ref{percentage_overlapping_node}, which is between 10\% to 60\% of the total CR nodes in the network.
In the other hand, the number of clusters heads is dependent on the network density and the CR node's transmission range as mentioned in Section~\ref{phaseI}.
The simulation in \cite{robust_clustering_arxiv} shows that the cluster heads are only 3.4\% to 20\% of the total CR nodes.

@@

\item \textbf{Robustness of the clusters against newly added PUs.}
We increase the number of PUs to challenge the non-singleton clusters, and count the number of the unclustered CR nodes.
This metric indicates the robustness of the clusters, \ie as to the clusters formed for a given CRN and spectrum availability, how many CR nodes can still be benefited from the clusters when the spectrum availability decreases.

@@

\subsubsection{CCs in Non-singleton Clusters}
\label{ccc_20}
%We first have a look at the average number of common channels per non-singleton cluster.
% shows the average number of the CCs of non-singleton clusters,
%as the singleton clusters (in other words unclustered nodes) don't execute any functionalities of clusters, which are has be discussed in Section~\ref{intro}.
From Figure~\ref{ccc_per_nonsingleton}, we can see the centralized schemes outperform the distributed schemes.
Among the distributed schemes, SOC achieves the most CCs.
The reason is, SOC is liable to group the neighboring CRs which share the most abundant spectrum together, no matter how many of them are there, thus the number of CC of the formed clusters is higher.
%We have discussed the flaw of this metric as it doesn't convey the number of unclustered CR nodes, in fact, 
On the other hand, SOC generates the most unclustered CRs.
As to the variants of ROSS, we notice that the greedy mechanism increases CCs in non-singleton clusters significantly.
%, this is due to the procedure that debatable nodes greedily look for better affiliation to improve the number of CCs.
%We also notice that the size control feature doesn't affect the number of CCs for both ROSS-DGA and ROSS-DFA.
%Size control mechanism converts the large clusters into small ones, but meanwhile clusters with the desired sizes have to be made when forming smaller clusters is possible.
%, which can also be observed in the large scale CRN in Section~\ref{largeScaleCRN}.

@@

\begin{figure*}[th]
\begin{multicols}{3}
    \includegraphics[width=\linewidth]{ccc_20.pdf}\par\caption{Average number of CCs of non-singleton clusters}\label{ccc_per_nonsingleton}
    \includegraphics[width=\linewidth]{cdf_clusterSize_20.pdf}\par\caption{Cumulative distribution of CRs residing in clusters with different sizes}\label{size_control}    
    \includegraphics[width=\linewidth]{survival_rate_20.pdf}\par\caption{Number of unclustered CRs with decreasing spectrum availability}\label{singleton_clusters}
\end{multicols}
\caption{Comparison between the distributed and centralized clustering schemes ($N$ = 20)}
\label{compare_dis_centralized}
\end{figure*}

@@

\subsubsection{Robustness of the formed clusters}
%When the number of PUs in CRN increases, or their operation becomes more intensive, some clusters don't seize any CCs any more, so that the cluster members and the cluster heads become unclustered, or singleton clusters.
%If there is no common channels available any more because of the new added PUs, the cluster is regarded as destroyed and the former cluster member CRs become unclustered CRs or in other words singleton clusters.
%We investigate the number of singleton clusters with primary users whose intensity of activities are varying.
%We will obtain two observations by examining this metric.
%First, we will know how many CR nodes are unclusterred by applying the robust clustering schemes in a given CRN.
In this part of simulation, we put PUs sequentially into CRN to decrease the available spectrum.
10 PUs are in the network in the beginning, then extra 19 batches of PUs are added sequentially, where each batch includes 5 PUs. 
%
Figure~\ref{singleton_clusters} shows certain clusters can not maintain and the number of unclustered CR nodes grows when the number of PUs increases.
The centralized scheme with desired size of 2 generates the most robust clusters, meanwhile, SOC results in the most vulnerable clusters.
The centralized scheme with desired size of 3 doesn't outperform the variants of ROSS, because pursuing cluster size prevents forming the the clusters with more CCs.
In contrary, the variants of ROSS generate some smaller clusters which are more likely to maintain when there are more PUs.

@@

\begin{figure}[ht!]
  \centering
  \includegraphics[width=.7\linewidth]{ccc_large_scale_color.pdf}
  \caption{Average number of CCs of non-singleton clusters}
  \label{ccc_large_scale}
\end{figure}

@@

\subsubsection{Robustness of the formed clusters}
We increase the primary users' activity by importing 20 batches of PUs sequentially in the CRN, where each batch includes 10 PUs.

@@

\begin{figure*}[t]
\begin{multicols}{3}
    \includegraphics[width=\linewidth]{survival_rate_100_edge50.pdf}\par\caption{100 CRs}\label{singleton_clusters_100}
    \includegraphics[width=\linewidth]{survival_rate_200_edge50.pdf}\par\caption{200 CRs}\label{singleton_clusters_200}
    \includegraphics[width=\linewidth]{survival_rate_300_edge50.pdf}\par\caption{300 CRs}\label{singleton_clusters_300}
\end{multicols}
%\caption{Percentage of CR nodes which are not included in any non-singleton clusters}
%\label{unclustered_100_200_300}
\end{figure*}

@@

\begin{figure*}[t]
\begin{multicols}{3}
    \includegraphics[width=\linewidth]{cdf_clusterSize_100.pdf}\par\caption{100 CRs, 30 PUs in network}\label{cdf_clusterSize_100}
    \includegraphics[width=\linewidth]{cdf_clusterSize_200.pdf}\par\caption{200 CRs, 30 PUs in network}\label{cdf_clusterSize_200}
    \includegraphics[width=\linewidth]{cdf_clusterSize_300.pdf}\par\caption{300 CRs, 30 PUs in network}\label{cdf_clusterSize_300}
\end{multicols}
%\caption{Cumulative distribution of CRs residing in clusters with different sizes}
%\label{cdf_100_200_300}
\end{figure*}
